---
title: "mypackage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mypackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Hello! welcome to mypackage, a demo package for DSA 5041. It consists of a 
custom t.test function and a few s3 methods for displaying the results more
cleanly. There is also a fun shiny server included!

The primary function is myttest, which creates a list of class "Rttest".
The other two functions, print.Rttest and plot.Rttest, serve to output the data
from myttest in a legible and useful way. 

```{r setup, warning=FALSE, message=FALSE}
library(mypackage)
```

Firstly, the main function myttest. Given two data sets, performs a t.test, 
either Student's, Welch, or Paired, depending on the data. 

Outputs a list of class Rttest. The list includes, in order, a data frame of x and y, 
a confidence interval of the t.test, the type of test used, 
p.value of the t.test, a rejection statement, and the alpha value used.

```{r myttest code, warning=FALSE, message=FALSE} 

myttest = function(x, y, alpha, paired) {

###Creates data frame from x and y
  df = data.frame(df.x = x, df.y = y)

### Determines which t.test to use based on equality of variance
  varx = var(x)
  vary = var(y)

  #If test are paired, runs the paired t.test
  if (paired == TRUE) {
    test_type = "Paired"
    t.vals = t.test(x,y, paired = TRUE, conf.level = 1 - alpha)
  }

  #If variances of x and y are equal and unpaired, runs a student's t.test
  else if (varx == vary) {
    test_type = "Student's"
    t.vals = t.test(x,y, var.equal = TRUE, conf.level = 1 - alpha)
  }

  #If variances are not equal and unpaired, runs welch test
  else {
    test_type = "Welch"
    t.vals = t.test(x,y, var.equal = FALSE, conf.level = 1 - alpha)
  }

### Determines if null hypothesis is rejected
  if (is.na(t.vals$p.value) == TRUE) {
    reject = "P-value is NaN, you might have an issue with your data."
  }
  else if(t.vals$p.value <= alpha) {
    reject = "T.test indicates that you should reject the Null."
  }
  else if (t.vals$p.value > alpha) {
    reject = "T.test indicates that you should *not* reject the Null."
  }

### Creates confidence interval
  CI = c(t.vals$conf.int[1],t.vals$conf.int[2])

### Creates P value
  p.val = t.vals$p.value

### Compiles everything into a list
  output_list = list(data.frame = df,
                     confidence.interval = CI,
                     test_type = test_type,
                     p.value = p.val,
                     reject = reject,
                     alpha = alpha)

### Sets custom class Rttest
  class(output_list) = "Rttest"

  output_list
}

```

Next, print.Rttest. This is a method of printing the results from myttest. It
returns a list of length 6: the kable table of x and y, 
a confidence interval of the t.test, the type of test used, 
the p.value of the t.test, a rejection statement, and the alpha value used.

```{r print.Rttest code, warning=FALSE, message=FALSE}

print.Rttest = function(x, ...) {
#kable is broken inside rmarkdown functions for some reason, so it outputs a raw html table. Sorry.
  print_output = list(kable.table = kable(x$data.frame), 
                      confidence.interval = x$confidence.interval, 
                      test_type = x$test_type, 
                      p.value = x$p.value,
                      alpha = x$alpha,
                      reject = x$reject)
  print_output
}


```

Lastly, plot.Rttest. This function defines the plot method for myttest. 
Depending on if the data is set as paired or not, the function will return one 
of two varieties of boxplots. For unpaired data, it returns side by side 
boxplots or a single boxplot of the difference in x and y for paired data. The 
paired data also includes the confidence interval on top of the graph.

```{r plot.Rttest code, warning=FALSE, message=FALSE}

plot.Rttest = function(x, ...) {
    obj = x.data.frame.df.x...x.data.frame.df.y = variable = value = 0

    ### creates plots for paired tests
    if (x$test_type == "Paired") {
      
      #uses purrr::map to round the CI list, for printing on the graph
      mapped = purrr::map(x[2], function(CI) round(CI,3))
      CI.list = c(as.character(mapped[[1]][1]), as.character(mapped[[1]][2]))
      confidence.interval.print = paste(CI.list, collapse = ", ")

      #creates a single plot of the difference of x and y
      paired_df = data.frame(x$data.frame$df.x - x$data.frame$df.y)
      plot = ggplot(paired_df, aes(y = x.data.frame.df.x...x.data.frame.df.y)) +
              geom_boxplot() +
              labs(x = "Difference of samples x and y", y = "values") +
              annotate("text", x = 0, y = 5, label = confidence.interval.print[1])
    }
    
    ### creates plots for unpaired tests
    else {
      #uses reshape2's melt to create the data frame for side-by-side plots
      long_df = reshape2::melt(x$data.frame) 

      plot = ggplot(long_df, aes(x = variable, y = value)) +
              geom_boxplot() +
              labs(x = "Samples x and y")
    }

    plot #outputs the plot
    }



```


And now, some examples for all the different cases. Warning, kable is kinda
broken for some reason; everywhere I looked had different reasons for why and 
different fixes, but nothing quite worked. Enjoy a bunch of raw html table data.

```{r function demo 1, warning=FALSE, message=FALSE}

set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans1=mypackage::myttest(x,y,alpha=0.05,paired=FALSE)
print(ans1)
plot(ans1)


```



```{r functions demo 2, warning=FALSE, message=FALSE}


set.seed(32); x=rnorm(30,mean=10,sd=5)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans2=mypackage::myttest(x,y,alpha=0.05,paired=FALSE)
print(ans2)
plot(ans2)


```



```{r functions demo 3, warning=FALSE, message=FALSE}

set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y = x+ rnorm(30, 5 ,4)
ans3=mypackage::myttest(x,y,alpha=0.05,paired=TRUE)
print(ans3)
plot(ans3)


```
